---
title: "Prepare CA WUI tract pop data"
output: html_notebook
---
```{r setup, warning=FALSE}
# loading libraries
library(tidyverse)
library(zoo)
library(foreign)
library(here)
library(purrr)
library(lubridate)
library(stringr)
```

- Clean workspace
```{r}
rm(list = ls())
```

- Read tract-level population file
```{r}
df_pop <- read_csv(here("data","pop_change_ca_tracts_2020_2022.csv"))
```

- Read WUI tract info
```{r}
df_wui <- read_csv(here("data","pop_trend_2010_2021_WUI_tracts_type.csv"))
```

- Join tract-level population file to WUI tract info

```{r}
df_wui_pop <- df_wui %>%
  left_join(df_pop, by = c("geo_id"="GEO_ID","county"="county")) %>%
  select(-county_pop_21, -tract_id_num)
```

- Obtain the number of tracts in each county of California

```{r}
tract_info <- df_pop %>%
  group_by(county) %>%
  summarise(tract_cnt = n())

```

- Add this info to `df_wui_pop`
```{r}
df_wui_pop <- df_wui_pop %>%
  left_join(tract_info, by = "county")
```

- Aggregate the tracts in WUI by county

```{r}
df_wui_pop_agg <- df_wui_pop %>%
  group_by(county) %>%
  summarise(cnt_incre_20_21 = sum(chg_20_21 > 0,na.rm = TRUE),
            cnt_incre_21_22 = sum(chg_21_22 > 0,na.rm = TRUE),
            cnt_incre_20_22 = sum(chg_20_22 > 0,na.rm = TRUE),
            cnt_decre_20_21 = sum(chg_20_21 < 0,na.rm = TRUE),
            cnt_decre_21_22 = sum(chg_21_22 < 0,na.rm = TRUE),
            cnt_decre_20_22 = sum(chg_20_22 < 0,na.rm = TRUE),
            total_tract = first(tract_cnt),
            pct_incre_20_21 = round(cnt_incre_20_21/total_tract*100,2),
            pct_incre_21_22 = round(cnt_incre_21_22/total_tract*100,2),
            pct_incre_20_22 = round(cnt_incre_20_22/total_tract*100,2),
            pct_decre_20_21 = round(cnt_decre_20_21/total_tract*100,2),
            pct_decre_21_22 = round(cnt_decre_21_22/total_tract*100,2),
            pct_decre_20_22 = round(cnt_decre_20_22/total_tract*100,2),
            .groups = 'drop')
```

- Aggregate the tracts in WUI by county for tracts inside significanty increased tracts

```{r}
df_wui_pop_agg_increased_tracts <- df_wui_pop %>%
  filter(mk_pop>0 & pval_mk < 0.05) %>%
  group_by(county) %>%
  summarise(cnt_incre_20_21 = sum(chg_20_21 > 0,na.rm = TRUE),
            cnt_incre_21_22 = sum(chg_21_22 > 0,na.rm = TRUE),
            cnt_incre_20_22 = sum(chg_20_22 > 0,na.rm = TRUE),
            cnt_decre_20_21 = sum(chg_20_21 < 0,na.rm = TRUE),
            cnt_decre_21_22 = sum(chg_21_22 < 0,na.rm = TRUE),
            cnt_decre_20_22 = sum(chg_20_22 < 0,na.rm = TRUE),
            total_tract = n(),
            pct_incre_20_21 = round(cnt_incre_20_21/total_tract*100,2),
            pct_incre_21_22 = round(cnt_incre_21_22/total_tract*100,2),
            pct_incre_20_22 = round(cnt_incre_20_22/total_tract*100,2),
            pct_decre_20_21 = round(cnt_decre_20_21/total_tract*100,2),
            pct_decre_21_22 = round(cnt_decre_21_22/total_tract*100,2),
            pct_decre_20_22 = round(cnt_decre_20_22/total_tract*100,2),
            .groups = 'drop')
```

- Save aggregated results for mapping

```{r}
write_csv(df_wui_pop_agg, here("data","wui_pop_change_cnty_2020_2022.csv"))
write_csv(df_wui_pop_agg_increased_tracts, 
          here("data","wui_sign_increase_pop_change_cnty_2020_2022.csv"))
write_csv(df_wui_pop, here("data","wui_pop_change_tract_2020_2022.csv"))
```